---
title: "1 - Preprocessing"
author: "Santiago García Sánchez"
format: html
editor: source
---

## Libraries

```{r, warning=F, message=FALSE}
library(readxl)
library(tidyverse)
library(naniar)
```

## Data loading

Download from [here](http://doi.org/10.5255/UKDA-SN-9174-3).

```{r}
datos <- read_excel(
  "childrens_pans_data_waves_1_10_holiday_2021_to_term_2025_safeguarded.xlsx",
  sheet = 2, col_types = "text" # Data imported as text
)
```

## Preprocessing

### Data selection

Only the school wave from 2025 is selected:

```{r}
datos <- datos |>
  filter(Year == 2025 & Period == "Term Wave")
```

Unnecesary columns are removed:

```{r}
datos$Interview_Date <- NULL
datos$CS_S_Child_Availability <- NULL
datos$CS_S_Child_Consent <- NULL
datos$CS_S_Disability_Consent <- NULL
datos$Wave <- NULL
datos$Period <- NULL
datos$Year <- NULL
```

Columns with 100% `NA` are removed:
```{r}
var_na <- miss_var_summary(datos) |>
  filter(pct_miss == 100) |>
  pull(variable)

datos <- datos |>
  select(!all_of(var_na))

rm(var_na)
```

Constants are removed:

```{r}
datos$CS_S_Q1_2 <- NULL
datos$CS_S_Q1_5 <- NULL
datos$CS_S_Q3_6 <- NULL
datos$CS_S_Q3_7 <- NULL
datos$CS_QX_1_13 <- NULL
datos$CS_QX_1_14 <- NULL
datos$CS_Q3_11 <- NULL
datos$CS_Q3_12 <- NULL
datos$CS_Q10_14 <- NULL
datos$CS_Q19_14 <- NULL
```

Columns that are combinations of other variables are removed:

```{r}
datos$CS_S_Q3_5 <- NULL
datos$CS_Q3_13 <- NULL
datos$CS_Q4_10 <- NULL
datos$CS_Q4_11 <- NULL
datos$CS_Q4_12 <- NULL
datos$CS_QX_1_12 <- NULL
datos$CS_Q10_15 <- NULL
datos$CS_Q10_16 <- NULL
```

### Transforming and recoding

Categorical variables are transformed back to numeric:

```{r, warning=FALSE}
datos$Weight_Percent <- as.numeric(datos$Weight_Percent)

datos <- datos |>
  mutate(
    across(
      starts_with("CS_S_Q2"),
      ~ replace_na(as.numeric(ifelse(. == "6 and over", "6", .)), 0)
    )
  )
```

Nominal variables are transformed into factors:

```{r}
# S (screening)
datos$CS_S_Region <- as.factor(datos$CS_S_Region)
datos$CS_S_SchoolFee <- as.factor(datos$CS_S_SchoolFee)

# Q (questions)
datos$CS_QX_3_B <- as.factor(datos$CS_QX_3_B)
```

Ordinal variables are transformed into factors:

```{r}
# S (screening)
datos$CS_S_SchoolTrip <- factor(datos$CS_S_SchoolTrip,
  levels = c(
    "Don’t know", "Never", "Once or twice",
    "Every term", "Once a month", "At least once a week or more"
  )
)

datos$CS_S_Income <- factor(datos$CS_S_Income,
  levels = c("Prefer not to say", "Dont know", "£0– £14,999", "£15,000–£19,999", "£20,000–£29,999", "£30,000–£39,999", "£40,000–£49,999", "£50,000+")
)

# Wellbeing
datos$CS_Wellbeing_happy <- ifelse(datos$CS_Wellbeing_happy == "0 Not at all", "0", datos$CS_Wellbeing_happy)
datos$CS_Wellbeing_happy <- ifelse(datos$CS_Wellbeing_happy == "10 Completely", "10", datos$CS_Wellbeing_happy)
datos$CS_Wellbeing_happy <- ifelse(datos$CS_Wellbeing_happy == "Prefer not to say", "5", datos$CS_Wellbeing_happy)
datos$CS_Wellbeing_happy <- as.numeric(datos$CS_Wellbeing_happy)

datos$CS_Wellbeing_anxious <- ifelse(datos$CS_Wellbeing_anxious == "0 Not at all", "0", datos$CS_Wellbeing_anxious)
datos$CS_Wellbeing_anxious <- ifelse(datos$CS_Wellbeing_anxious == "10 Completely", "10", datos$CS_Wellbeing_anxious)
datos$CS_Wellbeing_anxious <- ifelse(datos$CS_Wellbeing_anxious == "Prefer not to say", "5", datos$CS_Wellbeing_anxious)
datos$CS_Wellbeing_anxious <- as.numeric(datos$CS_Wellbeing_anxious)

datos$CS_Wellbeing_lonely <- factor(datos$CS_Wellbeing_lonely,
  levels = c("Prefer not to say", "Dont know", "Never", "Hardly ever", "Occasionally", "Some of the time", "Often / always")
)

# Q (Questions)
datos$CS_Q5A <- factor(datos$CS_Q5A,
  levels = c(
    "Prefer not to say", "Dont know", "Never", "Once or twice",
    "Most days", "Every day"
  )
)
datos$CS_Q5B <- factor(datos$CS_Q5B,
  levels = c(
    "Prefer not to say", "Dont know", "Never", "Once or twice",
    "Most days", "Every day"
  )
)
datos$CS_Q5C <- factor(datos$CS_Q5C,
  levels = c(
    "Prefer not to say", "Dont know", "Never", "Once or twice",
    "Most days", "Every day"
  )
)

datos$CS_Q6A <- factor(datos$CS_Q6A,
  levels = c(
    "Prefer not to say", "Dont know", "I don’t have a garden",
    "Never", "Once or twice", "Most days", "Every day"
  )
)
datos$CS_Q6B <- factor(datos$CS_Q6B,
  levels = c("Prefer not to say", "Dont know", "Never", "Once or twice", "Most days", "Every day")
)
datos$CS_Q6C <- factor(datos$CS_Q6C,
  levels = c("Prefer not to say", "Dont know", "Never", "Once or twice", "Most days", "Every day")
)

datos$CS_Q9 <- ifelse(datos$CS_Q9 == "1 Not at all important", "1", datos$CS_Q9)
datos$CS_Q9 <- ifelse(datos$CS_Q9 == "5 Very important", "5", datos$CS_Q9)
datos$CS_Q9 <- factor(datos$CS_Q9,
  levels = c("Dont know", "Prefer not to say", c(1:5))
)

datos$CS_Q11 <- factor(
  recode(datos$CS_Q11,
    A = 1, B = 2, C = 3,
    D = 4, E = 5, F = 6, G = 7
  ),
  levels = 1:7
)

datos$CS_Q12a <- factor(datos$CS_Q12a,
  levels = c(
    "Strongly disagree", "Disagree",
    "Neither agree nor disagree ", "Agree",
    "Strongly agree", "Completely agree"
  )
)
datos$CS_Q12b <- factor(datos$CS_Q12b,
  levels = c(
    "Completely disagree", "Strongly disagree",
    "Disagree", "Neither agree nor disagree",
    "Agree", "Strongly agree", "Completely agree"
  )
)
datos$CS_Q12c <- factor(datos$CS_Q12c,
  levels = c(
    "Completely disagree", "Strongly disagree",
    "Disagree", "Neither agree nor disagree",
    "Agree", "Strongly agree", "Completely agree"
  )
)
datos$CS_Q12e <- factor(datos$CS_Q12e,
  levels = c(
    "Completely disagree", "Strongly disagree",
    "Disagree", "Neither agree nor disagree",
    "Agree", "Strongly agree", "Completely agree"
  )
)

datos[paste0("CS_Q13_", LETTERS[1:5])] <- lapply(datos[paste0("CS_Q13_", LETTERS[1:5])], function(x) {
  x <- recode(x,
    "0 - Disagree a lot"  = "0",
    "10 - Agree a lot"    = "10"
  )
  factor(x, levels = c("Prefer not to say", "Don’t know", as.character(0:10)))
})

datos$CS_Q18 <- factor(datos$CS_Q18,
  levels = c("No", "Not sure", "Yes")
)

datos[paste0("CS_Q20_", 1:7)] <- lapply(datos[paste0("CS_Q20_", 1:7)], function(x) {
  x <- recode(x,
    "1 - Most important" = "1",
    "7 - Least important" = "7"
  )
  factor(x, levels = c("Prefer not to say", "Don’t know", as.character(0:7)))
})

datos$CS_QX_3_A <- factor(datos$CS_QX_3_A,
  levels = c("Mostly when I was not at school", "Equal time while I was at school, and not at school", "Mostly while I was at school")
)
datos$CS_QX_3_C <- factor(datos$CS_QX_3_C,
  levels = c("Prefer not to say", "Dont know", "Yes, I enjoy school less", "No, I enjoy school the same amount", "Yes, I enjoy school more")
)

datos[paste0("CS_QX_4_", LETTERS[1:4])] <- lapply(datos[paste0("CS_QX_4_", LETTERS[1:4])], function(x) {
  x <- recode(x,
    "0 - Disagree a lot"  = "0",
    "10 - Agree a lot"    = "10"
  )
  factor(x, levels = c("Prefer not to say", "Don’t know", as.character(0:10)))
})
```

Binary variables are turned into factors: 

```{r}
# S (screening)
datos$CS_S_Age_Band <- factor(datos$CS_S_Age_Band, levels = c("8 to 11", "12 to 15"))
# Gender is made binary
datos <- datos |>
  filter(CS_S_Gender %in% c("Female", "Male"))
datos$CS_S_Gender <- factor(datos$CS_S_Gender)

# Q (questions)
datos <- datos |>
  mutate(across(
    c(
      starts_with("CS_Q2_"), starts_with("CS_Q3_"),
      starts_with("CS_QX_1"), starts_with("CS_QY"),
      starts_with("CS_Q7"), starts_with("CS_Q8"),
      starts_with("CS_Q10"), starts_with("CS_Q11"),
      starts_with("CS_Q16"), starts_with("CS_Q17"),
      starts_with("CS_Q19")
    ),
    as.factor
  ))
```

### Creation of new variables

Computing households below the poverty line:

```{r}
# Calculation of income:
datos <- datos |>
  mutate(
    num = recode(CS_S_Income,
      "£0– £14,999" = 7500,
      "£15,000–£19,999" = 17500,
      "£20,000–£29,999" = 25000,
      "£30,000–£39,999" = 35000,
      "£40,000–£49,999" = 45000,
      "£50,000+" = 75000,
      .default = NA_real_
    )
  )

# Calculation of household size
datos$denom <- 1 +
  0.5 * replace_na(datos$CS_S_Q2_3, 0) +
  0.5 * replace_na(datos$CS_S_Q2_4, 0) +
  0.3 * replace_na(datos$CS_S_Q2_1, 0) +
  0.3 * replace_na(datos$CS_S_Q2_2, 0)


# Calculation of poverty
datos$ingresos <- datos$num / datos$denom

mediana <- 33800

datos$poverty <- ifelse(datos$ingresos < 0.6 * mediana, "Yes", "No")

# Temporary variables are removed
rm(mediana)
datos$num <- NULL
datos$denom <- NULL
datos$ingresos <- NULL

# Missing data is imputed as "Yes"
datos$poverty <- datos$poverty |>
  replace_na("Yes")
```

For question `CS_Q20`, new variables are created showing main incentives by person: 

```{r}
datos <- datos |>
  select(c(starts_with("CS_Q20_"), Respondent_ID)) |>
  pivot_longer(
    cols = starts_with("CS_Q20_"),
    names_to = "opcion",
    values_to = "rango"
  ) |>
  mutate(priority = paste0("Priority_", rango)) |>
  mutate(
    opcion = recode(opcion,
      "CS_Q20_1" = "Toilets nearby",
      "CS_Q20_2" = "Playground",
      "CS_Q20_3" = "Places to play sports",
      "CS_Q20_4" = "Safe spaces to cycle",
      "CS_Q20_5" = "Good weather",
      "CS_Q20_6" = "Clean air",
      "CS_Q20_7" = "People around"
    ),
    priority = factor(
      paste0("Priority_", rango),
      levels = paste0("Priority_", 1:7)
    )
  ) |>
  select(Respondent_ID, priority, opcion) |>
  pivot_wider(
    names_from = priority,
    values_from = opcion
  ) |>
  right_join(datos, by = "Respondent_ID") |>
  relocate(all_of(paste0("Priority_", 1:7)), .after = last_col())

# Mantener solo las primeras tres
datos <- datos |>
  select(-Priority_4:-Priority_7)
```

A numeric version of the `CS_Wellbeing_lonely` variable is created:

```{r}
datos <- datos |>
  mutate(
    CS_Wellbeing_lonely_num = recode(CS_Wellbeing_lonely,
      "Prefer not to say" = 6,
      "Dont know" = 6,
      "Never" = 2,
      "Hardly ever" = 4,
      "Occasionally" = 6,
      "Some of the time" = 8,
      "Often / always" = 10
    )
  )
```

School type (`CS_S_SchoolFee`) is simplified:

```{r}
datos <- datos |>
  mutate(
    CS_S_SchoolFee_2 = recode(CS_S_SchoolFee,
      "A State Run School" = "State Run",
      "A Fee Paying/Independent School" = "Fee Paying",
      .default = "Other type"
    )
  )
```

Number of means in which the child has heard about the Countryside Code:

```{r}
datos <- datos |>
  mutate(
    across(starts_with("CS_Q19"), ~ recode(.x, "Yes" = 1, "No" = 0)),
    means = rowSums(across(starts_with("CS_Q19")), na.rm = TRUE)
  )
```

Household access to green spaces:

```{r, warning=F}
datos <- datos |>
  mutate(
    CS_S_Q3_3 = recode(CS_S_Q3_3, "No" = 0, "Yes" = 1) |> as.numeric(),
    across(
      starts_with("CS_S_Q3_") & -all_of("CS_S_Q3_3"),
      ~ recode(.x, "No" = 0, "Yes" = 2) |> as.numeric()
    ),
    access = rowSums(pick(starts_with("CS_S_Q3_")), na.rm = TRUE),
    access = recode(
      access,
      `0` = "None",
      `1` = "One",
      .default = "Two or more"
    ),
    access = factor(access, levels = c("None", "One", "Two or more"))
  )
```

## Data control

Number of observations:

```{r}
length(unique(datos$Respondent_ID))
```
Variables with missing data:

```{r}
miss_var_summary(datos) |>
  filter(n_miss > 0)
```

## Data saving

Clean data is saved as a .rds file:

```{r}
saveRDS(datos, file = "clean_data.rds")
```
