---
title: "2 - Indicators"
author: "Santiago García Sánchez"
format: html
editor: source
---

## Libraries

```{r, warning=F, message=FALSE}
library(tidyverse)
library(dlookr)
library(FactoMineR)
library(factoextra)
```

## Data loading

```{r}
datos <- readRDS("clean_data.rds")
```

## Indicator 1: Time Spent Outdoors 

Recodificatión:
```{r}
# CS_Q3
datos$CS_Q3_3 <- as.numeric(recode(datos$CS_Q3_3, No = 0, Yes = 1))
datos <- datos |>
  mutate(across(
    starts_with("CS_Q3_") & !all_of("CS_Q3_3"),
    ~ recode(.x, "Yes" = 2, "No" = 0)
  ))

# CS_Q5A
datos <- datos |>
  mutate(CS_Q5A = as.numeric(recode(CS_Q5A,
    "Prefer not to say" = 0, "Dont know" = 0, "Never" = 0,
    "Once or twice" = 1, "Most days" = 2, "Every day" = 2
  )))

# CS_Q5B
datos <- datos |>
  mutate(CS_Q5B = as.numeric(recode(CS_Q5B,
    "Prefer not to say" = 0, "Dont know" = 0, "Never" = 0,
    "Once or twice" = 1, "Most days" = 2, "Every day" = 2
  )))
```

Computation:
```{r}
# Sum of the columns
datos <- datos |>
  mutate(
    Ind1 = rowSums(pick(starts_with("CS_Q3_"), CS_Q5A, CS_Q5B))
  ) |>
  # Rescaling
  mutate(
    Ind1 = as.numeric(10 * transform(Ind1, method = "minmax"))
  )
```

## Indicator 2: Green Spaces at School:

Recodification:
```{r}
# CS_QX_3_A
datos$CS_QX_3_A <- as.numeric(recode(datos$CS_QX_3_A,
  "Equal time while I was at school, and not at school" = 1,
  "Mostly when I was not at school" = 0,
  "Mostly while I was at school" = 2
))

# CS_QX_3_C
datos <- datos |>
  mutate(CS_QX_3_C = case_when(
    CS_QX_3_B == "Yes, there is more nature" & CS_QX_3_C == "Yes, I enjoy school more" ~ 1,
    CS_QX_3_B == "Yes, there is less nature" & CS_QX_3_C == "Yes, I enjoy school less" ~ 1,
    TRUE ~ 0
  ))

# CS_QX_3_B
datos$CS_QX_3_B <- as.numeric(recode(datos$CS_QX_3_B,
  "Yes, there is more nature" = 2,
  "No, there is the same amount" = 1,
  .default = 0
))


# CS_QX_4
datos <- datos |>
  mutate(across(
    starts_with("CS_QX_4"),
    ~ {
      x <- suppressWarnings(as.numeric(as.character(.)))
      case_when(
        x >= 5 & x <= 7 ~ 1,
        x >= 8 & x <= 10 ~ 2,
        TRUE ~ 0
      )
    }
  ))

# CS_QX_1
datos <- datos |>
  mutate(across(
    starts_with("CS_QX_1_"),
    ~ recode(.x, "Yes" = 1, "No" = 0)
  ))
```

Computation:
```{r}
# Sum of the columns
datos <- datos |>
  mutate(
    Ind2 = rowSums(pick(
      starts_with("CS_QX_3_"), starts_with("CS_QX_4"),
      starts_with("CS_QX_1")
    ), na.rm = TRUE)
  ) |>
  # Rescaling
  mutate(
    Ind2 = as.numeric(10 * transform(Ind2, method = "minmax"))
  )
```

## Indicator 3: Access to Green and Natural Places

Recodification:
```{r, warning=TRUE}
# CS_Q4
datos$CS_Q4_1 <- as.numeric(recode(datos$CS_Q4_1, No = 0, Yes = 1))
datos$CS_Q4_2 <- as.numeric(recode(datos$CS_Q4_2, No = 0, Yes = 1))
datos$CS_Q4_3 <- as.numeric(recode(datos$CS_Q4_3, No = 0, Yes = 2))
datos$CS_Q4_4 <- as.numeric(recode(datos$CS_Q4_4, No = 0, Yes = 2))
datos$CS_Q4_5 <- as.numeric(recode(datos$CS_Q4_5, No = 0, Yes = 2))
datos$CS_Q4_6 <- as.numeric(recode(datos$CS_Q4_6, No = 0, Yes = 2))
datos$CS_Q4_7 <- as.numeric(recode(datos$CS_Q4_7, No = 0, Yes = 2))
datos$CS_Q4_8 <- as.numeric(recode(datos$CS_Q4_8, No = 0, Yes = 2))
datos$CS_Q4_9 <- as.numeric(recode(datos$CS_Q4_9, No = 0, Yes = 2))
datos$CS_Q4_Other <- as.numeric(recode(datos$CS_Q4_Other, No = 0, Yes = 2))

# CS_Q6A
datos$CS_Q6A <- as.numeric(recode(datos$CS_Q6A,
  "Every day" = 2, "Most days" = 2, "Once or twice" = 1,
  .default = 0
))

# CS_Q6B
datos$CS_Q6B <- as.numeric(recode(datos$CS_Q6B,
  "Every day" = 2, "Most days" = 2, "Once or twice" = 1,
  .default = 0
))
```

Calculation:
```{r}
# Sum of the columns
datos <- datos |>
  mutate(
    Ind3 = rowSums(pick(starts_with("CS_Q4_"), CS_Q6A, CS_Q6B), na.rm = TRUE)
  ) |>
  # Rescaling
  mutate(
    Ind3 = as.numeric(10 * transform(Ind3, method = "minmax"))
  )
```

## Indicator 4: Quality of Local Natural Spaces

Recodification:
```{r, warning=FALSE}
# CS_Q13
datos <- datos |>
  mutate(across(
    starts_with("CS_Q13_"),
    ~ {
      x <- as.numeric(as.character(.))
      case_when(
        x >= 5 & x <= 7 ~ 1,
        x >= 8 & x <= 10 ~ 2,
        TRUE ~ 0
      )
    }
  ))
```

Calculation:
```{r}
# Sum of the columns
datos <- datos |>
  mutate(
    Ind4 = rowSums(pick(starts_with("CS_Q13_")), na.rm = TRUE)
  ) |>
  # Rescaling
  mutate(
    Ind4 = as.numeric(10 * transform(Ind4, method = "minmax"))
  )
```

## Indicator 5: Enjoyment of Nature and Nature Connection:

Recodification:
```{r, warning=FALSE}
# CS_Q11
datos$CS_Q11 <- case_when(
  as.numeric(datos$CS_Q11) %in% c(1, 2) ~ 0,
  as.numeric(datos$CS_Q11) %in% c(3, 4, 5) ~ 1,
  as.numeric(datos$CS_Q11) %in% c(6, 7) ~ 2,
  TRUE ~ 0
)

# CS_Q12a
datos$CS_Q12a <- as.numeric(recode(datos$CS_Q12a,
  "Completely agree" = 2, "Strongly agree" = 2, "Agree" = 1,
  .default = 0
))

# CS_Q5C
datos$CS_Q5C <- as.numeric(recode(datos$CS_Q5C,
  "Every day" = 2, "Most days" = 2, "Once or twice" = 1,
  .default = 0
))

# CS_Q6C
datos$CS_Q6C <- as.numeric(recode(datos$CS_Q6C,
  "Every day" = 2, "Most days" = 2, "Once or twice" = 1,
  .default = 0
))
```

Calculation:
```{r}
# Sum of the columns
datos <- datos |>
  mutate(
    Ind5 = rowSums(pick(CS_Q11, CS_Q12a, CS_Q5C, CS_Q6C), na.rm = TRUE)
  ) |>
  # Rescaling
  mutate(
    Ind5 = as.numeric(10 * transform(Ind5, method = "minmax"))
  )
```

## Indicator 6: Environmental Concern and Action

Recodification:
```{r, warning=FALSE}
# CS_Q9
datos$CS_Q9 <- case_when(
  as.numeric(datos$CS_Q9) %in% c(1, 2) ~ 0,
  as.numeric(datos$CS_Q9) %in% c(3, 4, 5) ~ 1,
  as.numeric(datos$CS_Q9) %in% c(6, 7) ~ 2,
  TRUE ~ 0
)

# CS_Q12b
datos$CS_Q12b <- as.numeric(recode(datos$CS_Q12b,
  "Completely agree" = 2, "Strongly agree" = 2, "Agree" = 1,
  .default = 0
))

# CS_Q12c
datos$CS_Q12c <- as.numeric(recode(datos$CS_Q12c,
  "Completely agree" = 2, "Strongly agree" = 2, "Agree" = 1,
  .default = 0
))

# CS_Q12e
datos$CS_Q12e <- as.numeric(recode(datos$CS_Q12e,
  "Completely agree" = 2, "Strongly agree" = 2, "Agree" = 1,
  .default = 0
))

# CS_Q10
datos <- datos |>
  mutate(across(
    starts_with("CS_Q10"),
    ~ recode(.x, "Yes" = 1, "No" = 0)
  ))

# CS_Q18
datos$CS_Q18 <- as.numeric(recode(datos$CS_Q18, "Yes" = 1, .default = 0))

# CS_Q16
datos$CS_Q16_11 <- as.numeric(recode(datos$CS_Q16_11, No = 0, Yes = 1))

datos <- datos |>
  mutate(across(
    starts_with("CS_Q16_") & !all_of("CS_Q16_11"),
    ~ recode(.x, "Yes" = 0, "No" = 1)
  ))

# CS_Q17
datos$CS_Q17_8 <- as.numeric(recode(datos$CS_Q17_8, No = 0, Yes = 1))

datos <- datos |>
  mutate(across(
    starts_with("CS_Q17_") & !all_of("CS_Q16_8"),
    ~ recode(.x, "Yes" = 0, "No" = 1)
  ))
```

Calculation:
```{r}
# Sum of the columns
datos <- datos |>
  mutate(
    Ind6 = rowSums(pick(
      CS_Q9, CS_Q12b, CS_Q12c, CS_Q12e, starts_with("CS_Q10"),
      starts_with("CS_Q19"), CS_Q18, starts_with("CS_Q16_"),
      starts_with("CS_Q17_")
    ), na.rm = TRUE)
  ) |>
  # Rescaling
  mutate(
    Ind6 = as.numeric(10 * transform(Ind6, method = "minmax"))
  )
```

## Indicator 7: Well-being

PCA is applied to well-being variables:
```{r}
vars_pca <- datos[, c(
  "CS_Wellbeing_happy", "CS_Wellbeing_anxious",
  "CS_Wellbeing_lonely_num"
)]

resultados <- PCA(
  vars_pca,
  graph = FALSE,
  row.w = datos$Weight_Percent
)
```

Sign is changed:
```{r}
resultados$ind$coord[, 1] <- resultados$ind$coord[, 1] * -1
resultados$var$coord[, 1] <- resultados$var$coord[, 1] * -1
resultados$var$cor[, 1] <- resultados$var$cor[, 1] * -1
resultados$svd$V[, 1] <- resultados$svd$V[, 1] * -1
```

Coordinates for each individual are obtained:
```{r}
coord <- get_pca_ind(resultados)$coord |>
  as_tibble() |>
  mutate(
    Dim.1 = as.numeric(10 * transform(Dim.1, method = "minmax")),
    Dim.2 = as.numeric(10 * transform(Dim.2, method = "minmax")),
    Dim.3 = as.numeric(10 * transform(Dim.3, method = "minmax"))
  )

datos$Ind7 <- coord$Dim.1
```

## Data saving

The dataset with the indicators is saved as a .rds file:

```{r}
saveRDS(datos, file = "data_indicators.rds")
```
